---
title: "coxph"
author: "yimin chen"
date: "2023-12-07"
output:
    pdf_document:
       latex_engine: xelatex
---



```{r echo = FALSE, message = FALSE, warning = FALSE}
#install.packages("gt")
library(knitr)
library(survival)
library(tidyverse)
library(ggplot2)
library(ggfortify)
library(survminer)
library(biostat3)
library(gt)
```
# In these data a subject changes exposure status from not bereaved to bereaved when his or her spouse dies. The first stage of the analysis therefore is to partition each follow–up into a record describing the period of follow-up pre–bereavement and (for subjects who were bereaved during the study) the period post–bereavement.

```{r, message = FALSE, warning = FALSE}
## Creating times relativ to spouse death (year=0)
brv2 <- mutate(brv,
               id=NULL,
               y_before_sp_dth =  as.numeric(doe -dosp) / 365.24,
               y_after_sp_dth = as.numeric(dox - dosp) / 365.24)

## Splitting at spouse death (year=0)
brvSplit <- survSplit(brv2, cut = 0, end="y_after_sp_dth", start="y_before_sp_dth", id="id",event="fail")

## Calculating risk times
brvSplit <- mutate(brvSplit,
                   t_sp_at_risk =   y_after_sp_dth - y_before_sp_dth,
                   brv = ifelse(y_after_sp_dth > 0, 1, 0))


```


```{r}
brvSplit2 <- mutate(brvSplit,
                    sex = as.factor(sex),
                    brv = as.factor(brv))
## Translate time scale from years from spouse death to ages
brvSplit3 <- brvSplit2 %>%
    mutate(age_sp_dth =  as.numeric(dosp - dob) / 365.24, # Age at spouse death
           age_start = age_sp_dth + y_before_sp_dth,      # Age at start of timeband
           age_end = age_sp_dth + y_after_sp_dth)         # Age at end of timeband

age_cat <- seq(70,100,5) # Split at these ages
brvSplit4 <- survSplit(brvSplit3, cut=age_cat, start="age_start", end="age_end", event="fail", zero = 0)

brvSplit4 <- mutate(brvSplit4,
                    t_at_risk = age_end- age_start, # Creating new time at risk
                    age = cut(age_end, age_cat))   # Creating age band category

## Calculate crude rates
survRate(Surv(t_at_risk, fail) ~ age, data=brvSplit4)
summary(coxph(Surv(age_start, age_end, fail) ~ brv,
              data = brvSplit4))
```

# also model these data using Cox regression. Provided we use the attained age as the time scale and split the data to obtain separate observations for the bereaved and non-bereaved person-time the following command will estimate the effect of bereavement adjusted for attained age.
```{r}
summary(coxph(Surv(age_start, age_end, fail) ~ brv + sex,
              data = brvSplit4))

summary(coxph(Surv(age_start, age_end, fail) ~ brv,
              data = brvSplit4))
```

# Use the Cox model to estimate the effect of bereavement separately for males and females and compare the estimates to those obtained using Poisson regression.
```{r}

summary(coxph(Surv(age_start, age_end, fail) ~ brv + sex,
              data = brvSplit4))
```

