---
title: "statistical method"
author: "yimin chen"
date: "2023-12-06"
output:
    pdf_document:
       latex_engine: xelatex
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
#install.packages("gt")
library(knitr)
library(survival)
library(tidyverse)
library(ggplot2)
library(ggfortify)
library(survminer)
library(biostat3)
library(gt)
```


```{r, message = FALSE, warning = FALSE}
library(biostat3)
## Creating times relativ to spouse death (year=0)
brv2 <- mutate(brv,
               id=NULL,
               y_before_sp_dth =  as.numeric(doe -dosp) / 365.24,
               y_after_sp_dth = as.numeric(dox - dosp) / 365.24)

## Splitting at spouse death (year=0)
brvSplit <- survSplit(brv2, cut = 0, end="y_after_sp_dth", start="y_before_sp_dth", id="id",event="fail")

## Calculating risk times
brvSplit <- mutate(brvSplit,
                   t_sp_at_risk =   y_after_sp_dth - y_before_sp_dth,
                   brv = ifelse(y_after_sp_dth > 0, 1, 0))


```


The mutate function from the dplyr package (assumed as it's not explicitly loaded but commonly used for such operations) is used to modify the brv data frame.
Two new columns are created: y_before_sp_dth and y_after_sp_dth. These represent the number of years before and after the death of a spouse (dosp), calculated by subtracting the date of the event of interest (doe or dox) from the date of the spouse's death and converting the difference into years (dividing by 365.24, the average number of days in a year accounting for leap years).


The survSplit function from the survival package is used to split the data into periods before and after the spouse's death. This is done by specifying a cut point at year 0 (the year of the spouse's death).
The function creates new observations in the dataset, splitting any observation that spans the time point 0 into two, one before and one after the spouse's death.
Further Data Transformation:

Another mutate function is used to calculate two new variables: t_sp_at_risk (the time at risk after the spouse's death, calculated as the difference between y_after_sp_dth and y_before_sp_dth) and brv (a binary indicator set to 1 if the event occurred after the spouse's death, otherwise 0).

```{r}
summary(brvSplit)
library(skimr)
skimr::skim(brvSplit)

```


```{r, message = FALSE, warning = FALSE}


surv_obj <- Surv(time = brv$dox-brv$doe, event = brv$fail)

# Fit Cox model
cox_model <- coxph(surv_obj ~., data = brv)


summary(cox_model)


cox.zph(cox_model)

```


```{r}


surv_object <- Surv(time = brvSplit$dox-brvSplit$doe, event = brvSplit$fail)

# Generate the life table using Kaplan-Meier estimate
life_table <- survfit(surv_object ~ 1)

# Print the life table
print(life_table)

life_table
plot(life_table, main = "Survival Curve", xlab = "Time", ylab = "Survival Probability")

```

male
```{r, message = FALSE, warning = FALSE}
lifetable1=lifetab2(Surv(time = brvSplit$dox-brvSplit$doe, brvSplit$fail==1) ~ 1, brvSplit[brvSplit$sex==1,], breaks = seq(0, 3000, by = 300))
print(lifetable1)
```



\newpage
female
```{r, message = FALSE, warning = FALSE}
lifetable2=lifetab2(Surv(time = brvSplit$dox-brvSplit$doe, brvSplit$fail==1) ~ 1, brvSplit[brvSplit$sex==2,], breaks = seq(0, 3000, by = 300))
print(lifetable2)
```



## KM and FH 
```{r , message = FALSE, warning = FALSE}
fit <- brvSplit%>%
  survfit(Surv(brvSplit$dox-brvSplit$doe, fail==1) ~ brv, data = .)

fit2 <- brvSplit %>%
  survfit(Surv(brvSplit$dox-brvSplit$doe, fail==0) ~ brv, data = ., type = "fleming")

splots <- list()

splots[[1]] <- ggsurvplot(fit, data = brvSplit, pval = TRUE, title = "Kaplan-Meier")
splots[[2]] <- ggsurvplot(fit2, data = brvSplit, pval = TRUE, title = "Fleming-Harrington")

arrange_ggsurvplots(splots, print = TRUE,
  ncol = 2, nrow = 1)
```


```{r, message = FALSE, warning = FALSE}
fit3 <- brvSplit%>%
  survfit(Surv(brvSplit$dox-brvSplit$doe, fail==1) ~ sex, data = .)

fit4 <- brvSplit %>%
  survfit(Surv(brvSplit$dox-brvSplit$doe, fail==0) ~ sex, data = ., type = "fleming")

splots <- list()

splots[[1]] <- ggsurvplot(fit3, data = brvSplit, pval = TRUE, title = "Kaplan-Meier")
splots[[2]] <- ggsurvplot(fit4, data = brvSplit, pval = TRUE, title = "Fleming-Harrington")

arrange_ggsurvplots(splots, print = TRUE,
  ncol = 2, nrow = 1)
```


```{r}
plot(fit3, conf.int = FALSE, col = "blue",
     xlab = "Time (Days)", ylab = "Estimated Survival Probability",
     main = "Comparison of S(t) between K-M and F-H methods")
lines(fit4, conf.int = FALSE, col = "red")
legend("topright", c("Kaplan-Meier", "Fleming-Harrington"),
       col = c("blue", "red"), lty = 1, bty = "n")
```



\newpage


```{r, message = FALSE, warning = FALSE}


km_fit <- survfit(surv_obj ~ group, data = brv)


ggsurvplot(km_fit, data = brv, pval = TRUE, conf.int = TRUE)

```

```{r, message = FALSE, warning = FALSE}
surv_obj <- Surv(time = brv$dox-brv$doe, event = brv$fail)

km_fit1 <- survfit(surv_obj ~ brv$sex)
ggsurvplot(km_fit1, data= brv,                  
           pval = TRUE,                
           conf.int = TRUE,           
           risk.table = TRUE,          
           ggtheme = theme_minimal(),  
           palette = "Dark",         
           main = "Kaplan-Meier Survival Curve", 
           xlab = "Days",              
           ylab = "Survival Probability") 
```

```{r, message = FALSE, warning = FALSE}
surv_obj <- Surv(time = brv$dox-brv$doe, event = brv$fail)

km_fit1 <- survfit(surv_obj ~ brv$group)
ggsurvplot(km_fit1, data= brv,                  
           pval = TRUE,                
           conf.int = TRUE,           
           risk.table = TRUE,          
           ggtheme = theme_minimal(),  
           palette = "Dark",         
           main = "Kaplan-Meier Survival Curve", 
           xlab = "Days",              
           ylab = "Survival Probability") 
```

```{r}
km_fit2 <- survfit(surv_obj ~brv$sex)


ggsurvplot(km_fit2, data=brv,                  
           pval = TRUE,                
           conf.int = TRUE,           
           risk.table = TRUE,          
           ggtheme = theme_minimal(),  
           palette = "Dark2",         
           main = "Kaplan-Meier Survival Curve", 
           xlab = "Time",              
           ylab = "Survival Probability") 
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
# Comparing survival curves between two or more groups
surv_diff <- survdiff(surv_obj ~ group, data = brv)

# Print test results
print(surv_diff)

```


## Log-rank test (death as event) comparing group
```{r}
surv_obj1 <- Surv(time = brv$dox-brv$doe, event = brv$fail)

log_rank_test <- survdiff(surv_obj1 ~ group, data = brv)

print(log_rank_test)
```

```{r}
surv_obj1 <- Surv(time = brv$dox-brv$doe, event = brv$fail)

log_rank_test2 <- survdiff(surv_obj1 ~ sex, data = brv)

print(log_rank_test2)
```

```{r}
# Creating the survival object
surv_obj <- Surv(time = brv$dox - brv$doe, event = brv$fail)

# Fit Cox model (specify variables or use '.' for all variables)
cox_model <- coxph(surv_obj ~ ., data = brv)
summary(cox_model)

# Check proportional hazards assumption
cox.zph(cox_model)
```


Kaplan-Meier Estimation:


$S(t) = \prod_{t_i \leq t} \left(1 - \frac{d_i}{n_i}\right)$


Where $S(t)$ is the survival probability at time $t_i$, $d_i$ is the number of events at time $t_i$, and $n_i$ is the number of subjects at risk at time $t_i$.


Cox Proportional Hazards Model:


$h(t) = h_0(t) \exp(\beta_1 \cdot \text{pspline(age)} + \beta_2 \cdot \text{size} + \beta_3 \cdot \text{grade} + \beta_4 \cdot \text{nodes} + \beta_5 \cdot \text{pgr} + \beta_6 \cdot \text{er} + \beta_7 \cdot \text{hormon} + \beta_8 \cdot \text{chemo})$


Where $h(t)$ is the hazard at time $t_i$, $h_0(t)$ is the baseline hazard, $\beta_1, \beta_2, ..., \beta_8$ are the coefficients for each covariate, which include age modeled with a penalized spline, tumor size, grade, number of positive lymph nodes, progesterone receptor levels, estrogen receptor levels, hormonal treatment, and chemotherapy, respectively.
