---
title: "statistical method"
author: "yimin chen"
date: "2023-12-06"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
library(survival)
library(tidyverse)
library(ggplot2)
library(ggfortify)
library(survminer)
library(biostat3)

```


```{r echo = FALSE, message = FALSE, warning = FALSE}
data = brv %>% 
  drop_na()
view(data)
```


```{r echo = FALSE, message = FALSE, warning = FALSE}


surv_obj <- Surv(time = brv$dox-brv$doe, event = brv$fail)

# Fit Cox model
cox_model <- coxph(surv_obj ~., data = brv)


summary(cox_model)


cox.zph(cox_model)

```
\newpage


```{r echo = FALSE, message = FALSE, warning = FALSE}


km_fit <- survfit(surv_obj ~ group, data = brv)


ggsurvplot(km_fit, data = brv, pval = TRUE, conf.int = TRUE)

```

```{r echo = FALSE, message = FALSE, warning = FALSE}
surv_obj <- Surv(time = brv$dox-brv$doe, event = brv$fail)

km_fit1 <- survfit(surv_obj ~ brv$sex)
ggsurvplot(km_fit1, data= brv,                  
           pval = TRUE,                
           conf.int = TRUE,           
           risk.table = TRUE,          
           ggtheme = theme_minimal(),  
           palette = "Dark",         
           main = "Kaplan-Meier Survival Curve", 
           xlab = "Days",              
           ylab = "Survival Probability") 
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
surv_obj <- Surv(time = brv$dox-brv$doe, event = brv$fail)

km_fit1 <- survfit(surv_obj ~ brv$group)
ggsurvplot(km_fit1, data= brv,                  
           pval = TRUE,                
           conf.int = TRUE,           
           risk.table = TRUE,          
           ggtheme = theme_minimal(),  
           palette = "Dark",         
           main = "Kaplan-Meier Survival Curve", 
           xlab = "Days",              
           ylab = "Survival Probability") 
```

```{r}
km_fit2 <- survfit(surv_obj ~brv$sex)


ggsurvplot(km_fit2, data=brv,                  
           pval = TRUE,                
           conf.int = TRUE,           
           risk.table = TRUE,          
           ggtheme = theme_minimal(),  
           palette = "Dark2",         
           main = "Kaplan-Meier Survival Curve", 
           xlab = "Time",              
           ylab = "Survival Probability") 
```

```{r echo = FALSE, message = FALSE, warning = FALSE}
# Comparing survival curves between two or more groups
surv_diff <- survdiff(surv_obj ~ group, data = brv)

# Print test results
print(surv_diff)

```


## Log-rank test (death as event) comparing group
```{r}
surv_obj1 <- Surv(time = brv$dox-brv$doe, event = brv$fail)

log_rank_test <- survdiff(surv_obj1 ~ group, data = brv)

print(log_rank_test)
```

```{r}
surv_obj1 <- Surv(time = brv$dox-brv$doe, event = brv$fail)

log_rank_test2 <- survdiff(surv_obj1 ~ sex, data = brv)

print(log_rank_test2)
```

```{r}
# Creating the survival object
surv_obj <- Surv(time = brv$dox - brv$doe, event = brv$fail)

# Fit Cox model (specify variables or use '.' for all variables)
cox_model <- coxph(surv_obj ~ ., data = brv)
summary(cox_model)

# Check proportional hazards assumption
cox.zph(cox_model)
```


Kaplan-Meier Estimation:


$S(t) = \prod_{t_i \leq t} \left(1 - \frac{d_i}{n_i}\right)$


Where $S(t)$ is the survival probability at time $t_i$, $d_i$ is the number of events at time $t_i$, and $n_i$ is the number of subjects at risk at time $t_i$.


Cox Proportional Hazards Model:


$h(t) = h_0(t) \exp(\beta_1 \cdot \text{pspline(age)} + \beta_2 \cdot \text{size} + \beta_3 \cdot \text{grade} + \beta_4 \cdot \text{nodes} + \beta_5 \cdot \text{pgr} + \beta_6 \cdot \text{er} + \beta_7 \cdot \text{hormon} + \beta_8 \cdot \text{chemo})$


Where $h(t)$ is the hazard at time $t_i$, $h_0(t)$ is the baseline hazard, $\beta_1, \beta_2, ..., \beta_8$ are the coefficients for each covariate, which include age modeled with a penalized spline, tumor size, grade, number of positive lymph nodes, progesterone receptor levels, estrogen receptor levels, hormonal treatment, and chemotherapy, respectively.
